### 2009-09-02 trunk/latex/paper1/contents.tex
adicionando referencias dos modelos.
parte de algoritmos.

@@

\subsection{ER (Erdos-Renyi)}

@@

Directed weighted networks with overlapping modules.

@@

We propose an extension to BCR model... which was proposed to model the links
between web pages.

The BCR+ model builds networks that respect a module-dependency network, that
is, a network that contains modules and allowed dependencies between modules. 
In the component dependency network, an edge between two vertices is only
allowed if there exist an edge between their respective modules.

First of all, the model generates a network that is a copy of the module
dependency network, each vertex belonging to its respective module. Then the
algorithm runs iteratively, and on each iteration it executes an operations that
modifies the network. It can add a new vertex together with an edge connecting it
to an old vertex, or it can add an edge between two existing vertices. The
choice of vertices, though, is not fully random. The probability that a
particular vertex v is choosen, P(v), is proportional to a function that involves the
in-degree or the out-degree of the node. We say that a vertex is chosen within a
set V according to a function f if

@@

out-neighbors(v) 
module(v)
neighbor-modules(v)

choose event with probabilities (p, q, r)
event 1:
  choose w within V according to f(x) = din + in-degree(x)
  add new vertex v to the module of w
  add edge from v to w
event 2:
  choose w within V according to f(x) = dout + out-degree(x)
  add new vertex v to the module of w
  add edge from w to v
event 3:
  choose v within V according to f(x) = dout + out-degree(x)
  choose case with probabilities (mu, 1 - mu)
  case 1: % out
    choose w within neighbor-modules(v) - neighbors(v) according to f(x) = din + in-degree(x)
    add edge from v to w
  case 2:
    choose w within module(v) - neighbors(v) according to f(x) = din + in-degree(x)
    add edge from v to w
end

picks a event that evolves
the network. Each event has an associated probability.

* Add a node with ongoing edge. With probability p, a new vertex is added to the network, together with an
edge from the new vertex to an existing vertex, chosen preferentially according
to .... Considering din = 0, this means that a vertex with indegree = 8 is
twice more likely to receive the edge than a vertex with indegree = 4. The
parameter din can be used to alleviate the handicap. (If din = 4, the first
vertex will only be 3/2 more likely to receive the edge). The new vertex is
put on the same cluster as the old vertex.

* Add a node with ingoing edge. With probability q, a new vertex is added to the network, together with an
edge from an existing vertex, chosen preferentially accordingly to dout +
outdegree, to the new vertex. The new vertex is put on the same cluster as the
old vertex. This case is similar to the previous case.

* Add an edge. With probability r, a new edge is added between two vertices, v and w. v is
chosen according to .... After that, with prob X, w is chosen from one of the
modules connected to v's module, else it's chosen from v's own module. In any
case, the exact choice is made according to ...

It's a growth model, that is, it can take as input an existing network and
evolve it.

TODO: Argument: why the imposed modules are natural modules?

@@

The CGW was proposed to model the evolution of software systems organized in
modules. It was proven formally to generate scale-free networks.

The CGW model is similar to BCR+ in which it is a growth model. The initial
network is composed of two vertices with a directed edge between, belonging
to the same module. The other M - 1 modules are initially empty.

Then, at each iteration, the algorithm executes one of the following events:

* With probability p1, one vertex is added to a randomly-chosen module, together
with e1 edges from p1 to vertices chosen according to module-based preferential
attachment.

* With probability

Accounts for the removal of edges. Growth model.

@@

>>>>

external edge...

@@

%\subsection{ER (Erdos-Renyi)}

@@

In common: the were proven formally to generate scale-free directed networks.

@@

We propose an extension to BCR model... which was proposed to model the links
between web pages.
\cite{Bollobas2003}

The BCR model aims to model the network of hyperlinks between web pages as a
directed graph without modules \cite{Bollobas2003}. It was proven analitically
to generate scale-free networks. We have developed an extension to this model,
which we will call BCR+, that adds the concept of module. The model accepts the
following parameters:

\begin{itemize}
\item number of vertices, $n$;
\item a graph, $G$;
\item three probabilities, $p$, $q$, and $r$, such as $p + q + r = 1$;
\item a probability, $mu$;
\item two real numbers, din and dout.
\end{itemize}

In a network with modules, one can define a MDG as a graph where the two
following conditions hold:
* each vertex represents a module in the original network;
* there is an edge from M1 to M2 in the MDG only if there is an edge from 
v1 to v2 in the original network, where v1 in M1 and v2 in M2.

The BCR+ model generates networks whose MDG is equal to G by adding vertices and
edges to an initial network until it reaches n vertices. The initial network is
isomorphic to G: it contains one vertex for each module and one edge for each
edge in G, connecting vertices whose corresponding modules are are connected.
Thus, the networks's MDG is equal to G from the very beginning.

After that, the algorithm consists of successive applications of one out of
three operations on the network: (1) adding a vertex with an outgoing edge; (2)
adding a vertex with an ingoing edge; (3) adding a edge between existing
vertices. The choice of the vertices that will be connected by a new edge,
although non-deterministic, is not fully random. The probability that a
particular vertex v is choosen, P(v), is proportional to a function of the
in-degree or of the out-degree of the node. We say that we choose a vertex
within a set V according to a function f if

@@

Before we present the algorithm in detail, using pseudo-code, we must explain
some definitions. V is the set of vertices in the network being generated. This
set grows as the algorithm is executed. Being v a vertex, we define the
following three functions:

* out-neighbors(v): the set of all vertices that are connected to v by an edge
starting in v.
* same-module(v): the set of all vertices that are in the same module as v (except
v itself).
* other\_modules(v): the set of all vertices that are in modules that are connected
to v's module (in the graph G) by an edge starting in v's module.

After the initial network is created, the algorithm proceeds as follows:

while the network has less than n vertices:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex v to w's module
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex v to w's module
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	choose a vertex w within (other\_modules(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within (same-module(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
  end

We can see that the probabilities p, q and r control how often each operation is
performed. Because the operation associated with probability r does not add any
vertices, greater values of r imply more edges in the resulting network. It is
easy to see that nodes connected to many ingoing edges are more likely to get
another ingoing edge, and the same reasoning can be applied to outgoing edges.
The parameter din can alleviate the handicap by providing a ``base in-degree''
that is applied to all vertices when computing the probabilities. Consider
two vertices, v1 with in-degree 4, and v2 with in-degree 8. If din = 0, v2 is
twice more likely to receive a new incoming edge; if, otherwise, din = 4, v2 is
only 3/2 more likely to receive the edge.

The parameter mu controls the proportion of edges between vertices in distinct
modules. Lower values of mu lend to networks that are more modular.

The BCR+ model is a growth model, meaning that the network is generated vertex
by vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down design methodologies CITE.

%* Add a node with ongoing edge. With probability p, a new vertex is added to the network, together with an
%edge from the new vertex to an existing vertex, chosen preferentially according
%to .... Considering din = 0, this means that a vertex with indegree = 8 is
%twice more likely to receive the edge than a vertex with indegree = 4. The
%parameter din can be used to alleviate the handicap. (If din = 4, the first
%vertex will only be 3/2 more likely to receive the edge). The new vertex is
%put on the same cluster as the old vertex.
%
%* Add a node with ingoing edge. With probability q, a new vertex is added to the network, together with an
%edge from an existing vertex, chosen preferentially accordingly to dout +
%outdegree, to the new vertex. The new vertex is put on the same cluster as the
%old vertex. This case is similar to the previous case.
%
%* Add an edge. With probability r, a new edge is added between two vertices, v and w. v is
%chosen according to .... After that, with prob X, w is chosen from one of the
%modules connected to v's module, else it's chosen from v's own module. In any
%case, the exact choice is made according to ...
%
%It's a growth model, that is, it can take as input an existing network and
%evolve it.
%
%TODO: Argument: why the imposed modules are natural modules?

@@

The CGW model was proposed to model the evolution of software systems organized in
modules. It was proven formally to generate scale-free networks. \cite{Chen2008}
It accepts the following parameters:

* n, m
* Probabilities p1, p2, p3, p4, summing 1
* Natural numbers e1, e2, e3, e4
* alpha

Just like BCR+, this is a growth model. Its initial network is composed of two
vertices belonging to the same module and a directed edge between them. The
remaining m - 1 modules are initially empty. Because the original implementation
of the model is not available, we describe in detail the algorithm we
implemented:

while the network has less than n vertices:
  choose operation (p1, p2, p3, p4)
  operation 1: (adding a vertex with e1 edges)
    create a new vertex v and add it to a randomly chosen module
    do e1 times:
      choose a vertex w according to mbpa(v)
      add an edge from v to w
  operation 2: (adding e2 edges)
    do e2 times:
      choose a vertex v randomly
      choose a vertex w according to mbpa(v)
      add edge from v to w
  operation 3: (rewiring e3 edges)
    do e3 times:
      choose a vertex v randomly
      choose an edge e randomly within edges that start in v
      choose w according to mbpa(v)
      remove e
      add an edge from v to w
  operation 3: (removing e4 edges)
    do e4 times:
      remove and randomly chosen edge

Unlike BCR+, this model does not allow constraints on the connection between
modules, however it accounts for the rewiring and the removal of edges.

@@

The LF model is a very flexible model that can generated weighted directed
networks with overlapping modules, that is, in which a vertex can belong to more
than one module. Unlike the previous models, this is not a growth model: all
vertices are generated at once and then the edges are added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at http://.
